### 2011-06-26 Text/Response_to_referee1.txt
Added bw hr figure, changed README to reflect submission date and added response to ref.
	modified:   README.md
	new file:   Text/Figures/hr_migration_bw.eps
	new file:   Text/Response_to_referee1.txt

This document contains the authors responses to the initial referee's report for ApJL submission LET26384. Several
changes have been made to the manuscript. The most significant changes are an expanded discussion section and an
additional figure, figure 4, depicting the aspect ratio of the disks at different times. Detailed responses to the
referee's comments follow below.

1. "Even though there are some interesting results in the paper, most of the major results on migration in
gravitationally unstable disks were already reported in Baruteau et al. (2011). There is therefore no need for fast
publication as a Letter, especially since there appears to be very little explanation of the results in the current
manuscript. I would encourage the authors to resubmit a more detailed version of this work to the main journal, in
particular focusing on the differences of their approach compared to that of Baruteau et al. (2011)."

We believe it is important to note that the paper referenced here by the referee had been accepted to the MNRAS only one
week before our submission. It appears that the article is currently "in press" and is only publicly available on
astro-ph. As such we feel it should not block the publication of this article to the ApJL. In addition, as outlined in
our extended comments in the discussion section, Baruteau et al. use very different physics in their protoplanetary disk
simulations. The simulations presented in our work are fully three dimensional and use realistic radiative physics,
while the Baruteau et al. simulations are two dimensional and use an idealized cooling prescription. It has been shown
by many authors that both 3D and realistic cooling are required to accurately characterize and measure GIs.

2. "The critical dimensionless parameter governing the linearity of planet-disk interactions is q/h^3, where q is the
mass ratio of the planet - central star, and h is the aspect ratio of the disk. A plot of h(r) would therefore be
helpful."

Disk aspect ratios for the end of both planet runs and the corresponding time in the fiducial run have been provided in
figure 4.

3. "Gap formation is governed by above parameter together with viscosity (Crida et al. 2006). In the current set-up,
gravitational and Reynolds stresses probably strongly depend on radius. A radial plot of the stresses would give some
insight in gap formation at various locations in the disk."

Although we would like to include such figures and relevant discussion, and will in our future more detailed article,
the constraints of the Letter format prevent us from doing so.

4. "If the planet does not open up a gap, it is subject to Type I migration. The migration rate can then be very
sensitive to background temperature and density profiles (Paardekooper et al. 2010). Plots of azimuthally averaged
surface density and temperature profiles would help to determine whether one would expect inward or outward Type I
migration."

In fact, one of the main points of the article is that distinctions such as type II, type II, and type III migration,
which depend on the presence and character of a gap in a laminar disk, need not apply in a GI active disk. Although the
overall migration rate is similar to that predicted by type I migration theory, the same mechanism is not acting here.

5. "The authors state that they do not resolve the width w of the horseshoe region of the planet. Again, the critical
parameter is q/h^3. For h=0.1, w should already be larger than the scale height. Even if h=0.2, w should be close to
h/2. Not resolving w therefore implies not resolving h, which would be quite bad. Are the authors sure they can not
resolve w?"

We have removed the statement which led the referee to this conclusion. What we intended to convey was the fact that we
lack sufficient resolution to accurately measure corotation torques, not that we could not resolve the horseshoe region.

6. "In the introduction, please note that Baruteau & Masset (2008) were actually the first to include self-gravity in
migration simulations (but in laminar disks only). It is the gravitationally unstable nature of the disk that is new in
the present study and in Baruteau et al. (2011)."

As this article deals exclusively with GI active disks, and we did not indicate that another article was the first to
include self-gravity we declined to cite Baruteau & Masset (2008).

7. "I find the 't=0 planet run' difficult to understand. The authors seem to claim that the presence of the planet
affects the development of GI. How can they justify these initial conditions? If the planet was formed by GI, the disk
should already be turbulent. If it formed in a laminar disk (or in a previous GI burst), it should already have migrated
inward."

We have include the following text to illuminate some of the possible scenarios: "This run investigates the effects of
an existing planet on the growth of GIs, which could occur if disk the disk undergoes episodic or localized
instabilities."

8. "I am not sure what the role of the ILR of the dominant spiral mode is in stopping the migration of the
planet. Intuitively, one might expect an ILR to push the planet inward. What does the torque balance look like?"

Again, space limitations prevent us from presenting full torque diagnostics, however they will be presented in our
follow-up article. As we say in the article, this is an area for further research.

9. "The interaction of the planet with all the GI modes appears to be quite stochastic. It seems to me that very little
can be concluded from a single run."

The simulations presented here are by no means comprehensive, but are intended to give an indication of what possible
effects GIs might have on migration, and vice-versa. We provide as much detail for the given runs as allowed in the
Letter format, including additional planet runs (as will be done in the follow-up article) is not feasible in the
allowed space.