### 2011-07-12 Text/migration_apjlsubmission.tex
Responded to second referee report, should be the final submission.
	new file:   Response_to_referee2.txt
	modified:   general.bib
	modified:   migration.tex
	modified:   migration_apjlsubmission.tex

%\documentclass[12pt,preprint]{aastex}

\documentclass[12pt,manuscript,authoryear]{aastex}

\usepackage{graphicx}
\citestyle{aa}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{grffile}

\begin{document}

\title{Migration of Gas Giant Planets in Gravitationally Unstable Disks}

\author{{\it Short Title: Gas Giant Migration ~~~~Article Type: Journal}}

\author{Scott Michael}
\affil{Department of Astronomy, Indiana University, Bloomington, IN 47405}
\email{scamicha@indiana.edu}

\and

\author{Richard H. Durisen}
\affil{Department of Astronomy, Indiana University, Bloomington, IN 47405}
\email{durisen@astro.indiana.edu}

\and

\author{Aaron C. Boley}
\affil{Department of Astronomy University of Florida, Gainesville, FL 32611}
\email{aaron.boley@gmail.com}

\begin{abstract} 
  Characterization of migration in gravitationally unstable disks is necessary to understand the fate of protoplanets
  formed by disk instability. As part of a larger study, we are using a 3D radiative hydrodynamics code to investigate
  how an embedded gas giant planet interacts with a gas disk that undergoes gravitational instabilities (GIs). This
  Letter presents results from simulations with a Jupiter-mass planet placed in orbit at 25 AU within a 0.14 $M_{\odot}$
  disk. The disk spans 5 to 40 AU around a 1 $M_{\odot}$ star and is initially marginally unstable. In one simulation,
  the planet is inserted prior to the eruption of GIs; in another, it is inserted only after the disk has settled into a
  quasi-steady GI-active state, where heating by GIs roughly balances radiative cooling. When the planet is present from
  the beginning, its own wake stimulates growth of a particular global mode with which it strongly interacts, and the
  planet plunges inward six AU in about 10$^3$ years. In both cases with embedded planets, there are times when the
  planet's radial motion is slow and varies in direction. At other times, when the planet appears to be interacting with
  strong spiral modes, migration both inward and outward can be relatively rapid, covering several AUs over hundreds of
  years. Migration in both cases appears to stall near the inner Lindblad resonance of a dominant low-order mode. Planet
  orbit eccentricities fluctuate rapidly between about 0.02 to 0.1 throughout the GI-active phases of the simulations.
\end{abstract}

\keywords{hydrodynamics --- instabilities --- planet-disk interactions --- planets and satellites: formation --- protoplanetary disks}

\label{firstpage}

\section{Introduction}

Migration of gas giant planets due to interactions with a circumstellar gas disk can play a major role in defining the
architecture of planetary systems. Work on migration \citep[see review by][]{papaloizou2007} has included gravitational
interaction of planets with both laminar and turbulent disks. However, radiative transport, detailed equations of state
(EOS), and the self-gravity of the gas disk have usually been ignored; the effects of a non-isothermal EOS have only
recently been included \citep[e.g.,][]{paardekooper2006,paardekooper2010,paardekooper2011}. Emerging studies, such as
\citet{baruteau2011} and the one reported here, are beginning to address some of these issues.  \citet{boss2005} and
\citet{mayer2004} examined radial migration of planet-mass fragments in gravitationally unstable disks, but their disks
were violently disrupted by fragmentation under conditions (radii $<$ 40 AU, disk-to-star mass ratios $M_d/M_s \sim
0.1$, and stellar mass $M_s \sim 1\;M_{\odot}$) where fragmentation may not actually occur
\citep{rafikov2005,rafikov2007,boley2006,boley2007b,boley2008,forgan2009,cai2010}. More recently, fragmentation into
clumps with gas giant or brown dwarf masses has been documented in numerical simulations of disks that are relatively
massive ($M_d/M_s \sim$ a few tenths) and spatially extended (outer radii $>$ 50 AU)
\citep{krumholz2007,stamatellos2007,stamatellos2009,boley2009,boley2010}, where fragmentation is expected from
semi-analytic arguments \citep[e.g.,][]{clarke2009,rafikov2009,dodson2009}. The fate of the clumps then depends in part
on their radial migration, which is a chaotic and messy affair in a fragmenting disk
\citep[e.g.,][]{boley2009,boley2010,vorobyov2010,boley2010b}. The occurrence of gravitational instabilities may be
episodic \citep[e.g.,][]{vorobyov2006,vorobyov2010b,zhu2010}. Clumps that survive and contract to the dimensions of
young planets can later find themselves in a disk that erupts again into GI activity. As the star/disk system evolves,
such a protoplanet may end up in a region of a GI-active disk where fragmentation does not occur.

To improve our understanding of how planets migrate in GI-active disks that are not fragmenting, we have begun a
systematic study to investigate the effects on both the disk and the planet by inserting a planet-mass object into disks
susceptible to GIs. Using techniques developed in earlier research by our group \citep{pickett2003,
  mejia2005,cai2008,boley2006,boley2007b,michael2010}, we can identify the dominant spiral waves in a simulation and
analyze how the waves interact with the planet's motion. Our goal is to determine both the effect of giant planets on
GIs and the effect of GIs on planet migration. Because GIs are sensitive to radiative physics, we use a well-tested
radiative scheme \citep{boley2007b} and realistic opacities \citep{dalessio2001}.

Section 2 below presents our numerical methods and initial conditions.  We describe the simulation results in \S3, and
discuss them in \S4.

\section{Computational Methodology}

\subsection{3-D Radiative Hydrodynamics}

The CHYMERA (Computational HYdrodynamics with MultiplE Radiation Algorithms) code \citep{boley2007b} is a second-order,
explicit, Eulerian scheme on a 3D cylindrical grid. The code uses a realistic equation of state for H$_2$
\citep{boley2007} and integrates an energy equation that includes $PdV$ work, net heating or cooling due to radiative
flux divergence, and heating by artificial bulk viscosity. Calculations are done on a uniform cylindrical grid with
reflection symmetry about the disk midplane and a grid size ($\varpi$,$\phi$,z) = (512,512,64).  The $z$-axis is the
rotation axis of the disk. The large number of azimuthal zones is necessary to resolve the planet's Hill sphere and the
planet's wake. These simulations utilize the radiative cooling scheme developed and tested in \citet{boley2007b}, where
the optically thick monochromatic flux in the $\varpi$ and $\phi$-directions is computed by flux-limited radiative
diffusion and where the radiative transport of energy in the $z$-direction is solved using a one-ray discrete ordinate
method in both optically thin and thick regions. Although the central star remains fixed at the grid center, we account
for acceleration of the reference frame by the planet and by the disk via indirect potentials, as in
\citet{michael2010}. The planet integration is done with a Verlet integrator \citep[e.g.,][]{hut1995}, and the indirect
potential terms are treated as in \citet{nelson2000a}. The Rosseland mean and Planck mean opacities and molecular
weights in our simulations are the same as those in \citet{boley2006,boley2007b}, except that we correct the
\citet{dalessio2001} gas mean molecular weight to $\mu$ = 2.33.

\subsection{Initial Model and Simulation Conditions}

The model disk, based on an equilibrium model from \citet{pickett2003}, orbits a 1 $M_{\odot}$ star and has a mass $M_d =
0.14 M_{\odot}$, inner and outer radii at 5 and 40 AU, and an initial surface density distribution $\Sigma \sim
\varpi^{-1/2}$. The time unit of one ORP (= outer rotation period) is defined as the rotation period of the initial disk
at $\varpi \approx$ 32 AU, or about 180 yr. The disk is initially located between radial grid zones 30 and 240 and is
close to isentropic, which results in a Toomre-$Q$ distribution with a marginally unstable \citep[see ][]{durisen2007}
minimum value of 1.38 at radial grid zone 161 (26.7 AU). The computational grid extends radially to 512 zones to
accommodate expansion of the outer disk once GIs become nonlinear. An outflow boundary condition is enforced at the
upper vertical grid boundary, the outer radial grid boundary, and an inner radial boundary at 2 AU. To seed
nonaxisymmetry, the density distribution is given an initial 0.01 \% random cell-to-cell perturbation.

This Letter presents three simulations. The first, which we call the {\sl fiducial run}, is simulated from time $t = 0$
to 21 ORP without a planet. In the second simulation, which we call the {\sl t = 0 planet run}, a one Jupiter-mass
($M_J$) planet with a roughly circular midplane orbit at 25 AU is included in the initial $t = 0$ ORP equilibrium
disk. This run investigates the effects of an existing planet on the growth of GIs, which could occur if disk the disk
undergoes episodic or localized instabilities. In the third simulation, which we call the {\sl t = 10 planet run}, a 1
$M_J$ planet with a similar orbit is inserted into the fiducial run at $t = 10.5$ ORP. We estimate semi-major axes $a$
and eccentricities $e$ from the planet's $\varpi(t)$ by using the maximum and minimum radii, $\varpi_{\rm max}$ and
$\varpi_{\rm min}$, over each complete orbit of the planet, where one orbit is $\Delta\phi=2\pi$. Specifically, $a =
(\varpi_{\rm max}+\varpi_{\rm min})/2$ and $e = (\varpi_{\rm max}-\varpi_{\rm min})/(\varpi_{\rm max}+\varpi_{\rm
  min})$. These parameters are intended to highlight the strong orbit-to-orbit variations in the motion of the planet.
A more refined estimate of the orbital elements is unnecessary because the potential of the disk, even if were
axisymmetric, will cause departures from Keplerian dynamics.

\section{Results}

{\it The Fiducial Run.} As in \citet{mejia2005}, the fiducial run exhibits four distinct phases: initial {\sl
  axisymmetric cooling}, the onset of GIs in a well-defined {\sl burst} of a few discrete low-order modes, a {\sl
  transition} to more complex nonaxisymmetric structure, and the {\sl asymptotic} establishment of quasi-steady GI
activity with an overall balance of heating and cooling. There is no fragmentation because this disk has a long cooling
time \citep{gammie2001,boley2006,boley2007b}.

{\it The $t = 0$ Planet Run.} Figure \ref{fig:Am} shows that inclusion of even this modest mass planet ($0.7 \%$ of
$M_d$) has a dramatic effect on the burst phase. Without a planet, when all modes grow from imposed noise, it takes 4
ORPs for coherent spiral modes to organize and grow to nonlinear amplitude. The growth is centered near the $Q$-minimum
at $26$ AU. Modes of cos($m\phi$) symmetry with $m = 4$ and 5 dominate. An $m = 3$ (three-armed) spiral also grows, but
somewhat more slowly, and lags the other modes in time. In the $t = 0$ planet run, the planet develops an organized wake
within about an ORP in which $m = 3$ is a dominant component ($\sim 5$ \% global density perturbation). This nonlinear
seeding causes $m = 3$ to dominate the GI burst, which also occurs $\sim$ 2 ORP earlier. Because of our initial
placement of the planet inside but close to the $Q$-minimum, the corotation radius (CR) of the $m = 3$ mode is fairly
close to that of the planet's orbit radius. When the triggered $m = 3$ mode becomes strongly nonlinear at about 3 ORP,
planet migration is significantly affected.

Figure \ref{fig:a} shows the evolution of the planet's radial position. From 0 to 2.5 ORP, the planet is torqued
primarily by its own wake and migrates inward.  Beginning at about 3 ORP, the planet interacts with the now nonlinear $m
= 3$ GI mode. At first, the planet gains angular momentum and moves outward, but, from 4 to 8 ORP, a time interval of
about 720 yr, the plant experiences a negative torque and plunges from 23 to 17 AU.  After $t = 8$ ORP, the main burst
is over, and the disk transitions into its asymptotic state, where modes of many $m$-values become comparably strong
(see Fig. \ref{fig:Am}). The planet's radial migration apparently stalls at about 16 to 17 AU. From an analysis of
periodicities present in the gas disk between 8 and 12 ORP, the planet lies a few AU inside the inner Lindblad resonance
of a strong $m = 2$ mode with CR at 29 AU in this early part of the asymptotic phase.

{\it The $t = 10$ Planet Run.} The motion of the planet in this case is more difficult to interpret. We see intervals of
fairly rapid inward or outward migration over several orbits, as well as times when radial migration appears to
stall. Between $t = 15$ and 19 ORP, the pattern of outward migration for 2 ORP followed by an inward plunge over the
next 2 ORP resembles the behavior in the $t = 0$ planet run during the burst, but, in this case, there is no distinct
transition between phases of GI activity. Animations of the evolution of the midplane density (available at
http://hdl.handle.net/2022/13304) show there is a complex interaction between the planet, its spiral wake, and the
global spiral arms caused by the instabilities. Analysis of images and periodicities suggests that, between 15 and 19
ORP, the planet may be in a mean motion resonance with $m = 2$ and 3 patterns, both of which have CR at 27 AU. Over this
time interval, the average orbit period of the planet is $\sim$ 0.50 ORP, while both the $m = 2$ and 3 spirals have a
pattern period of $\sim 0.74$ ORP, suggesting a 3:2 resonance. Both of these modes are present, at nearly the same
periodicities, in the fiducial run. So, in contrast with the $t = 0$ planet run, this planet's rapid phases of radial
migration are caused by interaction with strong GI modes that exist independent of the planet.

{\it Eccentricity.} Figure \ref{fig:e} shows the evolution of $e$ in the planet simulations. In both cases, the planets
were inserted with an approximate circular velocity computed by adding the interior disk mass to the stellar point
mass. The presence of GI activity in the $t = 10$ planet run immediately increases $e$ to $\sim 0.08$. In the $t = 0$
planet run, the modest initial $e$ decreases during the 1.5 ORP when it is migrating only due to its own wake, as
expected \citep{ward1998,ward2003,goldreich2004}.  However, once there are strong nonlinear interactions with GI modes,
$e$ jumps upward. In both runs, it appears that interaction with well-established GI activity leads to eccentricities
between 0.02 and 0.1 that vary in a chaotic way between these extremes on orbit period time scales. The magnitude of $e$
is roughly consistent with the ratio of the sound speed to orbital speed of the disk and the modest values for the Mach
numbers of the GI spirals \citep{boley2006a,boley2008}. In other words, the planet's orbit tends to have about as much
nonaxisymmetry as reflected in the pitch angles of the GI-induced spirals \citep{cossins2009}.

{\it Migration Relative to the Disk.} The disk surface density distribution does not vary much in the asymptotic phase,
so radial motion of the planet in the $t = 10$ planet run also represents radial migration relative to the gas. During
the burst \citep[e.g.,][]{boley2006}, the surface density of the disk changes dramatically. To verify that the planet is
really migrating relative to the background gas disk in the $t = 0$ run, we have compared the evolution of the planet's
radial position $\varpi_p(t)$ with the mass inside that radius, expressed in terms of $m_{\rm cyl}$, the fraction of the
disk mass interior to the planet. Between $t = 3.52$ and $3.89$ ORP, when $\varpi_p$ increases modestly, $m_{\rm cyl}$
increases from about 0.50 to 0.53; between 3.89 and 8.47 ORP, when $\varpi_p$ decreases dramatically, $m_{\rm cyl}$ also
decreases dramatically from 0.53 to 0.30. So, the planet is migrating significantly with respect to the disk mass even
while the inner part of the gas disk is moving inward during the burst. As a measure of the gas motion, $m_{\rm cyl}$ at
23 AU increases from 0.50 to 0.60 from $t =$ 3.89 to 8.47 ORP, corresponding to an average gas disk inflow rate of $\sim
10^{-5} M_{\odot}$ yr$^{-1}$.

\section{Discussion}

Even in non-self-gravitating disks, the presence of discrete radial structure is known to cause instabilities
\citep{li2000,li2001}. \citet{meschiari2008} demonstrated that a gap, presumably opened by an embedded planet, can cause
global modes to become unstable in self-gravitating disks that are otherwise stable against the development of GIs. Here
we are dealing with a planet that is not quite massive enough to open a gap, but its gravitational interaction with the
disk also has a strong effect on the growth and onset of GIs.

Figure \ref{fig:a} suggests that the overall radial drift during the first 10 ORP of the $t = 0$ planet run is not too
different from what one would get by extrapolating the migration due solely to the planet's wake during the first 2 or 3
ORP, where the average radial migration is about 4$\times10^{-3}$ AU yr$^{-1}$. Although there are no simple formulae
that are fully applicable to our radiatively cooled disk, using eq. (70) from \citet{tanaka2002} with parameters from
the simulated disk, we estimate 6$\times10^{-3}$ AU yr$^{-1}$ as an analytic expectation (see figure \ref{fig:hr} for
disk aspect ratios). While the interaction with a GI-active disk does not seem to result in a drastically different
overall migration rate, the direction and magnitude of migration in both simulations is quite variable and seems to
depend strongly on interactions of the planet with discrete modes in the disk. In $t = 10$ planet run, the overall
radial migration rate is similar but is even more erratic in direction than in the $t = 0$ planet run. The important
departures from laminar disk theory are threefold: the more erratic nature of migration, the importance of interactions
with discrete low-order modes, and stalling of migration near the inner edge of the strongest GI activity. The migration
is neither a random walk nor monotonic but has a chaotic character due to nonlinear interactions, similar to the erratic
migration seen in planet/disk simulations with Rossby vortex instabilities \citep{yu2010}. For all these reasons,
it will be important in future studies to explore the dependence of the planet's behavior on its mass and on its
placement relative to the site of GI-eruptions (for $t = 0$ type runs) and relative to the CRs of low-order GI modes
(for $t = 10$ type runs).

In both simulations, the inward migration appears to halt near 17 AU. This may simply be a coincidence, because the
overall average migration rates are similar, the starting positions are essentially the same, and the calculations have
a similar duration. We plan to extend simulations like these to much longer times in the future. However, the final
radial positions of both planets happen to be the location of the inner Lindblad resonance (ILR) for a strong m = 2 mode
present in both runs. This ILR is associated with a surface density enhancement in the disk. Near this enhancement, the
surface density gradient may be falling steeply enough to halt migration, a mechanism that can work in a laminar disk
\citep{paardekooper2009}.  Alternatively, at 17 AU, the level of GI activity is quite different interior to the planet
($Q \gtrsim 2$) and exterior to the planet ($Q \lesssim 2$), and this may affect the balance of the Lindblad torques
exerted by the disk on the planet. Whether the migration really does stall and what the mechanism may be are subjects
for further research.

Due to the global nature of the GIs and their strong interaction with the embedded planet, it is important that the gas
phase physics are as accurate as possible and detailed analysis of the interaction between the planet and global modes
is required. Of particular importance is the treatment of cooling in the disk. Although these simulations show some
similarities to \citet{baruteau2011}, their simulations employ an idealized cooling rate governed by the local dynamical
timescale while the simulations in this work use the more realistic radiative physics of
\citet{boley2007b}. Three-dimensional simulations with radiative physics are necessary for understanding the evolution
of spiral structure in vertically stratified disks because shocks and heating along spiral arms can change the disk's
vertical structure \citep[e.g.][]{boley2006a}, which affects the migration rate of planets.

{\it Summary.} The simulations in this Letter show that a planet of only 1$M_J$ placed near the $Q$-minimum in a
0.14$M_{\odot}$ disk can hasten the onset of gravitational instabilities and change the nature of the dominant modes
during the initial burst. Although the magnitude of radial migration (a few$\times10^{-3}$ AU yr$^{-1}$) is not too
different from that computed using the \citet{tanaka2002} isothermal disk formula based on the Lindblad torques, it can
fluctuate in sign on time scales on the order of the orbit period and can be strongly affected by gravitational
interactions of the planet with discrete GI-driven spiral modes, leading to phases of rapid radial motion. Planet orbit
eccentricities fluctuate between about 0.02 and 0.10. The simulations also suggest that planet migration may stall for
long periods near the ILRs of dominant global GI-modes.

\section{Acknowledgments}

The authors would like to thank F.C. Adams and J.C.B. Papaloizou for useful encouragement in this project. The research
was conducted with the support of NASA grants from the Origins of Solar Systems Program (NNG05GN11G and
NNX08AK36G). A.C.B. is currently a Sagan Postdoctoral Fellow at U. Florida. This material is based upon work supported
by the National Science Foundation under Grants No. OCI-0451237, OCI-0535258, OCI-0504075, and through TeraGrid resources provided by Indiana University, PSC, and NCSA.

\begin{thebibliography}{48}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{{Baruteau} {et~al.}(2011){Baruteau}, {Meru}, \&
  {Paardekooper}}]{baruteau2011}
{Baruteau}, C., {Meru}, F., \& {Paardekooper}, S.-J. 2011, ArXiv e-prints

\bibitem[{{Boley}(2009)}]{boley2009}
{Boley}, A.~C. 2009, ApJ, 695, L53

\bibitem[{{Boley} \& {Durisen}(2006)}]{boley2006a}
{Boley}, A.~C. \& {Durisen}, R.~H. 2006, ApJ, 641, 534

\bibitem[{{Boley} \& {Durisen}(2008)}]{boley2008}
---. 2008, ApJ, 685, 1193

\bibitem[{{Boley} \& {Durisen}(2010)}]{boley2010b}
---. 2010, \apj, 724, 618

\bibitem[{{Boley} {et~al.}(2007{\natexlab{a}}){Boley}, {Durisen}, {Nordlund},
  \& {Lord}}]{boley2007b}
{Boley}, A.~C., {Durisen}, R.~H., {Nordlund}, {\AA}., \& {Lord}, J.
  2007{\natexlab{a}}, ApJ, 665, 1254

\bibitem[{{Boley} {et~al.}(2007{\natexlab{b}}){Boley}, {Hartquist}, {Durisen},
  \& {Michael}}]{boley2007}
{Boley}, A.~C., {Hartquist}, T.~W., {Durisen}, R.~H., \& {Michael}, S.
  2007{\natexlab{b}}, ApJL, 656, L89

\bibitem[{{Boley} {et~al.}(2010){Boley}, {Hayfield}, {Mayer}, \&
  {Durisen}}]{boley2010}
{Boley}, A.~C., {Hayfield}, T., {Mayer}, L., \& {Durisen}, R.~H. 2010, Icarus,
  207, 509

\bibitem[{{Boley} {et~al.}(2006){Boley}, {Mej{\'{\i}}a}, {Durisen}, {Cai},
  {Pickett}, \& {D'Alessio}}]{boley2006}
{Boley}, A.~C., {Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Cai}, K., {Pickett},
  M.~K., \& {D'Alessio}, P. 2006, ApJ, 651, 517

\bibitem[{{Boss}(2005)}]{boss2005}
{Boss}, A.~P. 2005, ApJ, 629, 535

\bibitem[{{Cai} {et~al.}(2008){Cai}, {Durisen}, {Boley}, {Pickett}, \&
  {Mej{\'{\i}}a}}]{cai2008}
{Cai}, K., {Durisen}, R.~H., {Boley}, A.~C., {Pickett}, M.~K., \&
  {Mej{\'{\i}}a}, A.~C. 2008, ApJ, 673, 1138

\bibitem[{{Cai} {et~al.}(2010){Cai}, {Pickett}, {Durisen}, \&
  {Milne}}]{cai2010}
{Cai}, K., {Pickett}, M.~K., {Durisen}, R.~H., \& {Milne}, A.~M. 2010, \apjl,
  716, L176

\bibitem[{{Clarke}(2009)}]{clarke2009}
{Clarke}, C.~J. 2009, MNRAS, 612

\bibitem[{{Cossins} {et~al.}(2009){Cossins}, {Lodato}, \&
  {Clarke}}]{cossins2009}
{Cossins}, P., {Lodato}, G., \& {Clarke}, C.~J. 2009, \mnras, 393, 1157

\bibitem[{{D'Alessio} {et~al.}(2001){D'Alessio}, {Calvet}, \&
  {Hartmann}}]{dalessio2001}
{D'Alessio}, P., {Calvet}, N., \& {Hartmann}, L. 2001, ApJ, 553, 321

\bibitem[{{Dodson-Robinson} {et~al.}(2009){Dodson-Robinson}, {Veras}, {Ford},
  \& {Beichman}}]{dodson2009}
{Dodson-Robinson}, S.~E., {Veras}, D., {Ford}, E.~B., \& {Beichman}, C.~A.
  2009, ApJ, 707, 79

\bibitem[{{Durisen} {et~al.}(2007){Durisen}, {Boss}, {Mayer}, {Nelson},
  {Quinn}, \& {Rice}}]{durisen2007}
{Durisen}, R.~H., {Boss}, A.~P., {Mayer}, L., {Nelson}, A.~F., {Quinn}, T., \&
  {Rice}, W.~K.~M. 2007, Protostars and Planets V, 607

\bibitem[{{Forgan} {et~al.}(2009){Forgan}, {Rice}, {Stamatellos}, \&
  {Whitworth}}]{forgan2009}
{Forgan}, D., {Rice}, K., {Stamatellos}, D., \& {Whitworth}, A. 2009, MNRAS,
  394, 882

\bibitem[{{Gammie}(2001)}]{gammie2001}
{Gammie}, C.~F. 2001, ApJ, 553, 174

\bibitem[{{Goldreich} {et~al.}(2004){Goldreich}, {Lithwick}, \&
  {Sari}}]{goldreich2004}
{Goldreich}, P., {Lithwick}, Y., \& {Sari}, R. 2004, ApJ, 614, 497

\bibitem[{{Hut} {et~al.}(1995){Hut}, {Makino}, \& {McMillan}}]{hut1995}
{Hut}, P., {Makino}, J., \& {McMillan}, S. 1995, ApJL, 443, L93

\bibitem[{{Krumholz} {et~al.}(2007){Krumholz}, {Klein}, \&
  {McKee}}]{krumholz2007}
{Krumholz}, M.~R., {Klein}, R.~I., \& {McKee}, C.~F. 2007, \apj, 656, 959

\bibitem[{{Li} {et~al.}(2001){Li}, {Colgate}, {Wendroff}, \& {Liska}}]{li2001}
{Li}, H., {Colgate}, S.~A., {Wendroff}, B., \& {Liska}, R. 2001, \apj, 551, 874

\bibitem[{{Li} {et~al.}(2000){Li}, {Finn}, {Lovelace}, \& {Colgate}}]{li2000}
{Li}, H., {Finn}, J.~M., {Lovelace}, R.~V.~E., \& {Colgate}, S.~A. 2000, \apj,
  533, 1023

\bibitem[{{Mayer} {et~al.}(2004){Mayer}, {Quinn}, {Wadsley}, \&
  {Stadel}}]{mayer2004}
{Mayer}, L., {Quinn}, T., {Wadsley}, J., \& {Stadel}, J. 2004, ApJ, 609, 1045

\bibitem[{{Mej{\'{\i}}a} {et~al.}(2005){Mej{\'{\i}}a}, {Durisen}, {Pickett}, \&
  {Cai}}]{mejia2005}
{Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Pickett}, M.~K., \& {Cai}, K. 2005,
  ApJ, 619, 1098

\bibitem[{{Meschiari} \& {Laughlin}(2008)}]{meschiari2008}
{Meschiari}, S. \& {Laughlin}, G. 2008, \apjl, 679, L135

\bibitem[{{Michael} \& {Durisen}(2010)}]{michael2010}
{Michael}, S. \& {Durisen}, R.~H. 2010, \mnras, 406, 279

\bibitem[{{Nelson} {et~al.}(2000){Nelson}, {Benz}, \&
  {Ruzmaikina}}]{nelson2000a}
{Nelson}, A.~F., {Benz}, W., \& {Ruzmaikina}, T.~V. 2000, ApJ, 529, 357

\bibitem[{{Paardekooper} {et~al.}(2010){Paardekooper}, {Baruteau}, {Crida}, \&
  {Kley}}]{paardekooper2010}
{Paardekooper}, S., {Baruteau}, C., {Crida}, A., \& {Kley}, W. 2010, \mnras,
  401, 1950

\bibitem[{{Paardekooper} \& {Mellema}(2006)}]{paardekooper2006}
{Paardekooper}, S. \& {Mellema}, G. 2006, A\&A, 459, L17

\bibitem[{{Paardekooper} \& {Papaloizou}(2009)}]{paardekooper2009}
{Paardekooper}, S. \& {Papaloizou}, J.~C.~B. 2009, MNRAS, 394, 2283

\bibitem[{{Paardekooper} {et~al.}(2011){Paardekooper}, {Baruteau}, \&
  {Kley}}]{paardekooper2011}
{Paardekooper}, S.-J., {Baruteau}, C., \& {Kley}, W. 2011, \mnras, 410, 293

\bibitem[{{Papaloizou} {et~al.}(2007){Papaloizou}, {Nelson}, {Kley}, {Masset},
  \& {Artymowicz}}]{papaloizou2007}
{Papaloizou}, J.~C.~B., {Nelson}, R.~P., {Kley}, W., {Masset}, F.~S., \&
  {Artymowicz}, P. 2007, Protostars and Planets V, 655

\bibitem[{{Pickett} {et~al.}(2003){Pickett}, {Mej{\'{\i}}a}, {Durisen},
  {Cassen}, {Berry}, \& {Link}}]{pickett2003}
{Pickett}, B.~K., {Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Cassen}, P.~M.,
  {Berry}, D.~K., \& {Link}, R.~P. 2003, ApJ, 590, 1060

\bibitem[{{Rafikov}(2005)}]{rafikov2005}
{Rafikov}, R.~R. 2005, ApJL, 621, L69

\bibitem[{{Rafikov}(2007)}]{rafikov2007}
---. 2007, ApJ, 662, 642

\bibitem[{{Rafikov}(2009)}]{rafikov2009}
---. 2009, \apj, 704, 281

\bibitem[{{Stamatellos} {et~al.}(2007){Stamatellos}, {Hubber}, \&
  {Whitworth}}]{stamatellos2007}
{Stamatellos}, D., {Hubber}, D.~A., \& {Whitworth}, A.~P. 2007, \mnras, 382,
  L30

\bibitem[{{Stamatellos} \& {Whitworth}(2009)}]{stamatellos2009}
{Stamatellos}, D. \& {Whitworth}, A.~P. 2009, MNRAS, 392, 413

\bibitem[{{Tanaka} {et~al.}(2002){Tanaka}, {Takeuchi}, \& {Ward}}]{tanaka2002}
{Tanaka}, H., {Takeuchi}, T., \& {Ward}, W.~R. 2002, \apj, 565, 1257

\bibitem[{{Vorobyov} \& {Basu}(2006)}]{vorobyov2006}
{Vorobyov}, E.~I. \& {Basu}, S. 2006, ApJ, 650, 956

\bibitem[{{Vorobyov} \& {Basu}(2010{\natexlab{a}})}]{vorobyov2010}
---. 2010{\natexlab{a}}, \apjl, 714, L133

\bibitem[{{Vorobyov} \& {Basu}(2010{\natexlab{b}})}]{vorobyov2010b}
---. 2010{\natexlab{b}}, \apj, 719, 1896

\bibitem[{{Ward} \& {Hahn}(1998)}]{ward1998}
{Ward}, W.~R. \& {Hahn}, J.~M. 1998, AJ, 116, 489

\bibitem[{{Ward} \& {Hahn}(2003)}]{ward2003}
---. 2003, AJ, 125, 3389

\bibitem[{{Yu} {et~al.}(2010){Yu}, {Li}, {Li}, {Lubow}, \& {Lin}}]{yu2010}
{Yu}, C., {Li}, H., {Li}, S., {Lubow}, S.~H., \& {Lin}, D.~N.~C. 2010, ApJ,
  712, 198

\bibitem[{{Zhu} {et~al.}(2010){Zhu}, {Hartmann}, \& {Gammie}}]{zhu2010}
{Zhu}, Z., {Hartmann}, L., \& {Gammie}, C. 2010, ApJ, 713, 1143

\end{thebibliography}

\newpage

\begin{figure}[t]
\center
\includegraphics[width=12cm]{amt.eps}
\caption{(Color figure available in the online version) Global amplitudes of nonaxisymmetric density perturbations $A_m$
  as a function of time for individual cos($m\phi$) perturbations. The formula for computing $A_m$ is equation (15) of
  \citep{boley2006}. Here the integrals are done only over the disk outside 15 AU to suppress contributions from edge
  effects in the inner disk. The top and bottom panels correspond to the fiducial run (no planet) and the $t = 0$ planet
  run, respectively.}
\label{fig:Am}
\end{figure}

\begin{figure}[t]
\center
\includegraphics[width=12cm]{planeta.eps}
\caption{Plot of the planet's radial position as a function of time $\varpi_p(t)$ for the two planet simulations. The
  dashed-diamond and solid-star lines (colored red and black in the online version) indicate the curves ($\varpi(t)$),
  points ($a$), and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively. The
  symbols are the approximate semi-major axes $a$ computed for each 0 to 2$\phi$ change in azimuth of the planet as
  explained in the text.}
\label{fig:a}
\end{figure}

\begin{figure}[t]
\center
\includegraphics[width=12cm]{planete.eps}
\caption{Plot of the eccentricities $e$, computed for each 0 to 2$\phi$ change in azimuth of the planet as explained in
  the text, as a function of time. The diamonds and stars (colored red and black in the online version) indicate the
  points ($e$) and horizontal scales ($t$) that correspond to the $t = 0$ and $t = 10$ planet runs, respectively.}
\label{fig:e}
\end{figure}

\begin{figure}[t]
\center
\includegraphics[width=12cm]{hr_migration.eps}
\caption{(Color figure available in the online version) Plots of the disk aspect ratio $h/r$ as determined by a decrease
  from midplane density by a factor of $e^2$ for the fiducial and $t=0$ runs (black) at $10.5$ ORP and the fiducial and
  $t=10$ runs (light gray) at $20.5$ ORP.}
\label{fig:hr}
\end{figure}

%\begin{figure}[t]
%\includegraphics{../Figures/name_of_figure.eps}
%\caption{here's the caption}
%\label{fig:name}
%\end{figure}

\label{lastpage}

\end{document}